---
title: "Assignment 1 - Language development in autistic and neurotypical children"
output: html_document
group members: "Victoria Engberg Lowe, Vlada Caraman, Natasha Becker Bertelsen" 
date: "12/10-2022"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 1  - Language development in autistic and neurotypical children

```{r Load packages}
pacman::p_load(tidyverse, 
       glue,
       data.table,
       dplyr,
       moments,
       tidybayes,
       ggplot2,
       ggridges,
       plyr,
       ellipse,
       brms,
       cowplot,
       viridis,
       gridExtra,
       tidyr,
       loo)
```

```{r Simulate data}
set.seed(1000)

# Defining variables 
n <- 30
visits <- 6
mu_asd <- log(1.5)
sigma_asd <- log(1.5)-log(1.5-0.5)
mu_td <- log(1.5)
sigma_td <- log(1.5)-log(1.5-0.3)
mu_visit_asd <- 0.1
sigma_visit_asd <- 0.03 
mu_visit_td <- 0.2
sigma_visit_td <- 0.017
error <- 0.1 

d <- tibble(expand.grid(ID = seq(n), Diagnosis = c("ASD", "TD"), Visit = seq(visits)))

#changing id (so that every id has diff diagnosis)
d <- d %>% 
mutate(ID=ifelse(Diagnosis =="ASD", ID+(n*2), ID))

#generate individual intercept (MLU at v1) and idv slope (change by visit)
d <- d %>% mutate(
  IndividualIntercept = NA,
  IndividualSlope = NA
)

for (i in seq(d$ID)) {
  d$IndividualIntercept[d$ID == i & d$Diagnosis == "ASD"] <- rnorm(1, mu_asd, sigma_asd)
  d$IndividualIntercept[d$ID == i & d$Diagnosis == "TD"] <- rnorm(1, mu_td, sigma_td)
   d$IndividualSlope[d$ID == i & d$Diagnosis == "ASD"] <- rnorm(1, mu_visit_asd, sigma_visit_asd)
     d$IndividualSlope[d$ID == i & d$Diagnosis == "TD"] <- rnorm(1, mu_visit_td, sigma_visit_td)
}

#calculate mlu per each datapoint
for (i in seq(nrow(d))) {
  d$MLU[i] <- exp(rnorm(1,(d$IndividualIntercept[i]+
                          d$IndividualSlope[i]
                           *(d$Visit[i]-1)),
                        error))
}

#checking data
glimpse(d)

ggplot(d, aes(Visit, MLU, color = Diagnosis, group = ID)) +
  theme_bw()+
  geom_point()+
  geom_line(alpha=0.3)

```

```{r Define formula}

MLU_f1 <- bf(MLU ~ 0 + Diagnosis + Diagnosis:Visit + (1 + Visit|ID)) 

get_prior(MLU_f1, d, family = "lognormal")

```

```{r Define the prior}
# Define the prior 
MLU_p1 <- c(
  prior(normal(0, 0.2), class = b), 
  prior(normal(0.4, 0.1), class = b, coef = "DiagnosisASD"), 
  prior(normal(0.4, 0.1), class = b, coef = "DiagnosisTD"), 
  prior(normal(0, 0.1), class = sd, coef = Intercept, group = ID),
  prior(normal(0, 0.05), class = sd, coef = Visit, group = ID),
  prior(lkj(3), class = "cor")
)

```

```{r Prior predictive checks}
MLU_m1_prior <- brm(
  MLU_f1,
  data = d,
  family = lognormal,
  prior = MLU_p1,
  sample_prior = "only",
  chains = 2,
  cores = 2,
  backend = "cmdstanr",
    threads = threading(2),
  control = list(
    adapt_delta = 0.99,
    max_treedepth = 20
    
    
  )
)

pp_check(MLU_m1_prior, ndraws = 100) +
  labs(title = "Prior-predictive check for Model 1")

```

```{r Fit the model and prior-posterior checks}
MLU_m1_posterior <- brm(
    MLU_f1,
    data = d,
    family = lognormal,
    prior = MLU_p1,
    sample_prior = T,
    cores = 2, 
    chains = 2, 
    backend = "cmdstanr",
    threads = threading(2),
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 20
    )
  )

```

```{r  Model quality checks}
# Posterior predictive checks
pp_check(MLU_m1_posterior, ndraws = 100) +
  labs(title = "Posterior-predictive check Model 1")

plot(conditional_effects(MLU_m1_posterior))

# Prior-posterior update checks
posterior_m1 <- as_draws_df(MLU_m1_posterior)

# Intercept ASD update check
Intercept_ASD <- ggplot(posterior_m1) +
  geom_density(aes(prior_b_DiagnosisASD), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(b_DiagnosisASD), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab('Intercept ASD') +
  theme_classic()

# Intercept TD update check
Intercept_TD <- ggplot(posterior_m1) +
  geom_density(aes(prior_b_DiagnosisTD), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(b_DiagnosisTD), fill = "#FC4E07", 
               color = "black", alpha = 0.6) +
  xlab('Intercept TD') +
  theme_classic()

# Slope ASD update check
Slope_ASD <- ggplot(posterior_m1) +
  geom_density(aes(`prior_b_DiagnosisASD:Visit`), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(`b_DiagnosisASD:Visit`), fill = "#FC4E07", 
               color = "black", alpha = 0.6) +
  xlab('Slope ASD') +
  theme_classic()

# Slope TD update check
Slope_TD <- ggplot(posterior_m1) +
  geom_density(aes(`prior_b_DiagnosisTD:Visit`), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(`b_DiagnosisTD:Visit`), fill = "#FC4E07", 
               color = "black", alpha = 0.6) +
  xlab('Slope TD') +
  theme_classic()

# Overall standard deviation for ID (population level)
posterior_m1$prior_sd_ID__Intercept

SD_ID <- ggplot(posterior_m1) +
  geom_density(aes(prior_sd_ID__Intercept), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(sd_ID__Intercept), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab('SD ID') +
  theme_classic()

# Overall standard deviation for Visit (population level)
SD_Visit <- ggplot(posterior_m1) +
  geom_density(aes(prior_sd_ID__Visit), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(sd_ID__Visit), fill = "#FC4E07", 
               color = "black", alpha = 0.6) +
  xlab('SD Visit') +
  theme_classic()

# Correlation between intercept and visit 
Cor <- ggplot(posterior_m1) +
  geom_density(aes(prior_cor_ID, fill = "steelblue",
                   color = "black", alpha = 0.6)) +
  geom_density(aes(cor_ID__Intercept__Visit), fill = "#FC4E07", 
               color = "black", alpha = 0.6) +
  xlab('Correlation ID and Visit') +
  theme_classic()

# Sigma
sigma <- ggplot(posterior_m1) +
  geom_density(aes(prior_sigma), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(sigma), fill = "#FC4E07", 
               color = "black", alpha = 0.6) +
  xlab('Sigma') +
  theme_classic()

grid.arrange(Intercept_ASD, Intercept_TD, SD_ID, Slope_ASD, Slope_TD, SD_Visit, sigma, Cor, nrow = 4)

# Prior-posterior update check on the intercepts
p1 <- ggplot(posterior_m1) +
  geom_histogram(aes(prior_b_DiagnosisASD),
                 fill = "red", color = "red", alpha = 0.3, bins = 50) +
  geom_histogram(aes(b_DiagnosisASD),
                 fill = "green", color = "green", alpha = 0.3, bins = 50) +
  geom_histogram(aes(b_DiagnosisTD),
                 fill = "blue", color = "blue", alpha = 0.3, bins = 50) +
  theme_bw() +
  xlab("Prior-posterior update check on the intercepts")

# Prior-posterior update check on the variability of the intercept
p2 <- ggplot(posterior_m1) + 
  geom_histogram(aes(prior_sd_ID__Intercept),
                 fill = "red", color = "red", alpha = 0.3, bins = 50) +
  geom_histogram(aes(sd_ID__Intercept),
                 fill = "green", color = "green", alpha = 0.3, bins = 50) +
  theme_bw() + 
  xlab("Prior-posterior update check on the variability of the intercept")

# Prior-posterior update check on slope 
p3 <- ggplot(posterior_m1) +
  geom_histogram(aes(`prior_b_DiagnosisASD:Visit`),
                 fill = "red", color = "red", alpha = 0.3, bins = 50) +
  geom_histogram(aes(`b_DiagnosisASD:Visit`),
                 fill = "green", color = "green", alpha = 0.3, bins = 50) +
  geom_histogram(aes(`b_DiagnosisTD:Visit`),
                 fill = "blue", color = "blue", alpha = 0.3, bins = 50) +
  theme_bw() +
  xlab("Prior-posterior update check on the slopes")

# Prior-posterior update check on the variability of the slope
p4 <- ggplot(posterior_m1) + 
  geom_histogram(aes(prior_sd_ID__Visit),
                 fill = "red", color = "red", alpha = 0.3, bins = 50) +
  geom_histogram(aes(sd_ID__Visit),
                 fill = "green", color = "green", alpha = 0.3, bins = 50) +
  theme_bw() + 
  xlab("Prior-posterior update check on the variability of the slope")

# Prior-posterior update check on the correlation
p5 <- ggplot(posterior_m1) + 
  geom_histogram(aes(prior_cor_ID),
                 fill = "red", color = "red", alpha = 0.3, bins = 50) +
  geom_histogram(aes(cor_ID__Intercept__Visit),
                 fill = "green", color = "green", alpha = 0.3, bins = 50) +
  theme_bw() + 
  xlab("Prior-posterior update check on the correlation")

grid.arrange(p1,p2,p3,p4,p5)

```

```{r Model quality checks}
# Rhat and Effective samples 
summary(MLU_m1_posterior)

# The Rhat values provide information about the convergence of the algorithm. Rhat values close to 1 suggest that the model has converged. In this case, the  Rhat values are close to 1, and therefore the model has converged. 

# The Bulk_ESS and Tail_ESS (effective sample size (ESS)) capture the sampling efficiency in the bulk and tails of the distribution. 

```
Once the pipeline is in place, loop through different sample sizes to assess how much data you would need to collect. N.B. for inspiration on how to set this up, check the tutorials by Kurz that are linked in the syllabus.
```{r Power Analysis}
fit <- brm(
    MLU_f1,
    data = d,
    family = lognormal,
    prior = MLU_p1,
    sample_prior = T,
    cores = 2, 
    chains = 2, 
    backend = "cmdstanr",
    threads = threading(2),
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 20
    ),
    seed = 1
  )

# Creating a function for the simulation
sim_d_and_fit <- function(seed, n){
  
  set.seed(seed)
  
  visits <- 6
  mu_asd <- log(1.5)
  sigma_asd <- log(1.5)-log(1.5-0.5)
  mu_td <- log(1.5)
  sigma_td <- log(1.5)-log(1.5-0.3)
  mu_visit_asd <- 0.1
  sigma_visit_asd <- 0.03 
  mu_visit_td <- 0.2
  sigma_visit_td <- 0.017
  error <- 0.1 

  d <- 
    tibble(expand.grid(ID = seq(n), Diagnosis = c("ASD", "TD"), Visit = seq(visits))) %>% 
    mutate(IndividualIntercept = NA,
           IndividualSlope = NA)
  
  d <- d %>% 
    mutate(ID=ifelse(Diagnosis =="ASD", ID+(n*2), ID))

  for (i in seq(d$ID)) {
    d$IndividualIntercept[d$ID == i & d$Diagnosis == "ASD"] <- rnorm(1, mu_asd, sigma_asd)
    d$IndividualIntercept[d$ID == i & d$Diagnosis == "TD"] <- rnorm(1, mu_td, sigma_td)
    d$IndividualSlope[d$ID == i & d$Diagnosis == "ASD"] <- rnorm(1, mu_visit_asd, sigma_visit_asd)
    d$IndividualSlope[d$ID == i & d$Diagnosis == "TD"] <- rnorm(1, mu_visit_td, sigma_visit_td)
}

  for (i in seq(nrow(d))) {
    d$MLU[i] <- exp(rnorm(1,(d$IndividualIntercept[i]+
                               d$IndividualSlope[i]*(d$Visit[i]-1)),error))
  }
}

# Producing one simulation
test <- sim_d_and_fit(100,30)

# Iterating simulation 100 times 
n_sim <- 100

t1 <- Sys.time() # tracking the time

s <- #large nested df
  tibble(seed = 1:n_sim) %>% 
  mutate(d = purrr::map(seed, sim_d_and_fit, n = 30)) %>% 
  mutate(fit = purrr::map2(d, seed, ~update(fit, newdata = .x, seed = .y)))

t2 <- Sys.time()
t2-t1
head(s)
```

```{r}
# POWER ANALYSIS 3 TRY
steszt <- s %>% 
 mutate("betas" = purrr::map(fit, ~ as_draws_df(.) %>%  
                        data.frame() %>%  
                        tibble::rownames_to_column("parameters"))) %>% 
select(-d,-fit) %>% 
unnest("betas") %>%  
select(seed, b_DiagnosisASD.Visit, b_DiagnosisTD.Visit)

# SLOPE POSTERIORS
lab_colors <- c("TD" = "blue", "ASD" = "red") 
steszt2<-steszt %>% 
  dplyr::group_by(seed) %>% 
  dplyr::summarise(mean(b_DiagnosisASD.Visit),
            min(b_DiagnosisASD.Visit),
            max(b_DiagnosisASD.Visit),
            mean(b_DiagnosisTD.Visit),
            min(b_DiagnosisTD.Visit),
            max(b_DiagnosisTD.Visit)) %>% 

  mutate(power = ifelse( `min(b_DiagnosisTD.Visit)`- `max(b_DiagnosisASD.Visit)` > 0,1,0)) 

steszt2 %>% 
  ggplot(aes(x=seed)) +
  geom_pointrange(fatten = 1/2,aes(y = `mean(b_DiagnosisASD.Visit)`, ymin = `min(b_DiagnosisASD.Visit)`, ymax = `max(b_DiagnosisASD.Visit)`, color = "ASD"))+
  geom_pointrange(fatten = 1/2, aes(y = `mean(b_DiagnosisTD.Visit)`, ymin = `min(b_DiagnosisTD.Visit)`, ymax = `max(b_DiagnosisTD.Visit)`, color = "TD"))+
  labs(x = "seed (i.e., simulation index)",
       y = " ",
       color = "Legend",
       subtitle = stringr::str_glue("30 participants from each group, power = {mean(steszt2$power)}"))+
  scale_color_manual(values = lab_colors)

```

# Part 2 - Strong in the Bayesian ken, you are now ready to analyse the actual data
```{r}
# Loading in the real data
df <- read_csv("clinical_data.csv")

# Modifying the dataframe
df <- df %>% 
  dplyr::rename(MLU = Child_Mlu)

# Describing the sample
## Overview of columns and variable type
glimpse(df)

## Calculating means, min and max
### age
df %>% 
  dplyr::group_by(Visit) %>% 
  dplyr::summarize(Mean_age = mean(Age, na.rm = T))

### The number of female participants (55 in visit 1)
df %>% 
  dplyr::group_by(Visit) %>%   
  dplyr::filter(Gender == "Female") %>% 
  dplyr::summarize(n())

### The number of male participants (11 in visit 1)
df %>% 
  dplyr::group_by(Visit) %>%   
  dplyr::filter(Gender == "Male") %>% 
  dplyr::summarize(n())

### cognitive abilities
df %>% 
  dplyr::group_by(Diagnosis) %>% 
  dplyr::summarize(mean_non_verb = mean(Non_verbal_iq_1, na.rm = T),
                   mean_verb = mean(Verbal_iq_1, na.rm = T),
                   mean_soc = mean(Socialization_1, na.rm = T))
```
```{r Apply model to empirical data}
# Applying our model to empirical data
df_model_1 <- brm(
    MLU_f1,
    data = df,
    prior = MLU_p1,
    #family = lognormal,
    sample_prior = T,
    cores = 2, 
    chains = 2, 
    backend = "cmdstanr",
    threads = threading(2),
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 20
    )
  )

pp_check(df_model_1, ndraws = 100) + 
  labs(title = "Posterior-update check with real data")

plot(conditional_effects(df_model_1))

ggplot(df, aes(Visit, MLU, color = Diagnosis, group = ID)) +
  theme_bw()+
  geom_point()+
  geom_line(alpha=0.3)

summary(df_model_1)
```

```{r Model comparison}
set.seed(800)
# Scaling the predictors
df <- df %>% mutate_at(c("Mom_mlu",
                         "Verbal_iq_1",
                         "Socialization_1"), ~(scale(.) %>% as.vector))

# Defining new models 
MLU_f2 <- bf(MLU ~ 0 + Diagnosis + Diagnosis:Visit + Diagnosis:Mom_mlu + (1 + Visit|ID)) 
MLU_f3 <- bf(MLU ~ 0 + Diagnosis + Diagnosis:Visit + Diagnosis:Verbal_iq_1 + (1 + Visit|ID)) 
MLU_f4 <- bf(MLU ~ 0 + Diagnosis + Diagnosis:Visit + Diagnosis:Socialization_1 + (1 + Visit|ID))

# Set priors 
MLU_p2 <- c(
  prior(normal(0, 0.2), class = b), 
  prior(normal(0.4, 0.1), class = b, coef = "DiagnosisASD"), 
  prior(normal(0.4, 0.1), class = b, coef = "DiagnosisTD"), 
  prior(normal(0, 0.1), class = sd, coef = Intercept, group = ID),
  prior(normal(0, 0.05), class = sd, coef = Visit, group = ID),
  prior(lkj(3), class = "cor")
)

# Prior-predictive check model 2
df_model_2 <- brm(
    MLU_f2,
    data = df,
    prior = MLU_p2,
    sample_prior = "only",
    cores = 2, 
    chains = 2, 
    backend = "cmdstanr",
    threads = threading(2),
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 20
    )
  )

m2 <- pp_check(df_model_2, ndraws = 100) + 
  labs(title = "Prior-predictive check model 2")

# Prior-predictive check model 3
df_model_3 <- brm(
    MLU_f3,
    data = df,
    prior = MLU_p2,
    sample_prior = "only",
    cores = 2, 
    chains = 2, 
    backend = "cmdstanr",
    threads = threading(2),
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 20
    )
  )

m3 <- pp_check(df_model_3, ndraws = 100) + 
  labs(title = "Prior-predictive check model 3")

# Prior-predictive check model 4
df_model_4 <- brm(
    MLU_f4,
    data = df,
    prior = MLU_p2,
    sample_prior = "only",
    cores = 2, 
    chains = 2, 
    backend = "cmdstanr",
    threads = threading(2),
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 20
    )
  )

m4 <- pp_check(df_model_4, ndraws = 100) + 
  labs(title = "Prior-predictive check model 4")

grid.arrange(m2,m3,m4)

# Posterior-predictive check model 2
df_post_2 <- brm(
    MLU_f2,
    data = df,
    prior = MLU_p2,
    sample_prior = T,
    cores = 2, 
    chains = 2, 
    backend = "cmdstanr",
    threads = threading(2),
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 20
    )
  )

mpost2 <- pp_check(df_post_2, ndraws = 100) + 
  labs(title = "Posterior-predictive check model 2")

# Prior-predictive check model 3
df_post_3 <- brm(
    MLU_f3,
    data = df,
    prior = MLU_p2,
    sample_prior = T,
    cores = 2, 
    chains = 2, 
    backend = "cmdstanr",
    threads = threading(2),
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 20
    )
  )

mpost3 <- pp_check(df_post_3, ndraws = 100) + 
  labs(title = "Posterior-predictive check model 3")

# Prior-predictive check model 4
df_post_4 <- brm(
    MLU_f4,
    data = df,
    prior = MLU_p2,
    sample_prior = T,
    cores = 2, 
    chains = 2, 
    backend = "cmdstanr",
    threads = threading(2),
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 20
    )
  )

mpost4 <- pp_check(df_post_4, ndraws = 100) + 
  labs(title = "Posterior-predictive check model 4")

grid.arrange(mpost2,mpost3,mpost4)

```

```{r Posterior predictive update checks model 2}
# Prior-posterior update checks
posterior_m2 <- as_draws_df(df_post_2)
variables(posterior_m2)

# Intercept ASD update check
Intercept_ASD2 <- ggplot(posterior_m2) +
  geom_density(aes(prior_b_DiagnosisASD), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(b_DiagnosisASD), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab('Intercept ASD') +
  theme_classic() +
  labs(title = "Intercept ASD")

# Intercept TD update check
Intercept_TD2 <- ggplot(posterior_m2) +
  geom_density(aes(prior_b_DiagnosisTD), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(b_DiagnosisTD), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab('Intercept TD') +
  theme_classic() +
  labs(title = "Intercept TD")

# Slope ASD visit update check
Slope_ASD2a <- ggplot(posterior_m2) +
  geom_density(aes(`prior_b_DiagnosisASD:Visit`), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(`b_DiagnosisASD:Visit`), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab('Slope Visit ASD') +
  theme_classic() +
  labs(title = "Slope Visit ASD")

# Slope TD visit update check
Slope_TD2a <- ggplot(posterior_m2) +
  geom_density(aes(`prior_b_DiagnosisTD:Visit`), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(`b_DiagnosisTD:Visit`), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab('Slope Visit TD') +
  theme_classic() +
  labs(title = "Slope Visit TD")

# Slope ASD mom_mlu update check
Slope_ASD2b <- ggplot(posterior_m2) +
  geom_density(aes(`prior_b_DiagnosisASD:Mom_mlu`), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(`b_DiagnosisASD:Mom_mlu`), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab('Slope Mom_mlu ASD') +
  theme_classic() +
  labs(title = "Slope Mom_mlu ASD")

# Slope TD mom_mlu update check
Slope_TD2b <- ggplot(posterior_m2) +
  geom_density(aes(`prior_b_DiagnosisTD:Mom_mlu`), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(`b_DiagnosisTD:Mom_mlu`), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab('Slope Mom_mlu TD') +
  theme_classic() +
  labs(title = "Slope Mom_mlu TD")

# SD intercept update check
SD_intercept2 <- ggplot(posterior_m2) +
  geom_density(aes(prior_sd_ID__Intercept), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(sd_ID__Intercept), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab("SD intercept") +
  theme_classic() +
  labs(title = "SD intercept")

# SD visit update check
SD_visit2 <- ggplot(posterior_m2) +
  geom_density(aes(prior_sd_ID__Visit), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(sd_ID__Visit), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab("SD visit") +
  theme_classic() +
  labs(title = "SD visit")

# sigmaupdate check
Sigma2 <- ggplot(posterior_m2) +
  geom_density(aes(sigma), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(sigma), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab("Sigma") +
  theme_classic() +
  labs(title = "Sigma")

grid.arrange(Intercept_ASD2, Intercept_TD2, Slope_ASD2a, Slope_TD2a, Slope_ASD2b, Slope_TD2b, SD_intercept2, SD_visit2, Sigma2, ncol = 4)

```
```{r Posterior predictive update checks model 3}
# Prior-posterior update checks
posterior_m3 <- as_draws_df(df_post_3)
variables(posterior_m3)

# Intercept ASD update check
Intercept_ASD3 <- ggplot(posterior_m3) +
  geom_density(aes(prior_b_DiagnosisASD), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(b_DiagnosisASD), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab('Intercept ASD') +
  theme_classic() +
  labs(title = "Intercept ASD")

# Intercept TD update check
Intercept_TD3 <- ggplot(posterior_m3) +
  geom_density(aes(prior_b_DiagnosisTD), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(b_DiagnosisTD), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab('Intercept TD') +
  theme_classic() +
  labs(title = "Intercept TD")

# Slope ASD visit update check
Slope_ASD3a <- ggplot(posterior_m3) +
  geom_density(aes(`prior_b_DiagnosisASD:Visit`), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(`b_DiagnosisASD:Visit`), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab('Slope Visit ASD') +
  theme_classic() +
  labs(title = "Slope Visit ASD")

# Slope TD visit update check
Slope_TD3a <- ggplot(posterior_m3) +
  geom_density(aes(`prior_b_DiagnosisTD:Visit`), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(`b_DiagnosisTD:Visit`), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab('Slope Visit TD') +
  theme_classic() +
  labs(title = "Slope Visit TD")

# Slope ASD Verbal_iq_1 update check
Slope_ASD3b <- ggplot(posterior_m3) +
  geom_density(aes(`prior_b_DiagnosisASD:Verbal_iq_1`), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(`b_DiagnosisASD:Verbal_iq_1`), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab('Slope Verbal_iq_1 ASD') +
  theme_classic() +
  labs(title = "Slope Verbal_iq_1 ASD")

# Slope TD Verbal_iq_1 update check
Slope_TD3b <- ggplot(posterior_m3) +
  geom_density(aes(`prior_b_DiagnosisTD:Verbal_iq_1`), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(`b_DiagnosisTD:Verbal_iq_1`), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab('Slope Verbal_iq_1 TD') +
  theme_classic() +
  labs(title = "Slope Verbal_iq_1 TD")

# SD intercept update check
SD_intercept3 <- ggplot(posterior_m3) +
  geom_density(aes(prior_sd_ID__Intercept), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(sd_ID__Intercept), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab("SD intercept") +
  theme_classic() +
  labs(title = "SD intercept")

# SD visit update check
SD_visit3 <- ggplot(posterior_m3) +
  geom_density(aes(prior_sd_ID__Visit), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(sd_ID__Visit), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab("SD visit") +
  theme_classic() +
  labs(title = "SD visit")

# sigmaupdate check
Sigma3 <- ggplot(posterior_m3) +
  geom_density(aes(sigma), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(sigma), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab("Sigma") +
  theme_classic() +
  labs(title = "Sigma")

grid.arrange(Intercept_ASD3, Intercept_TD3, Slope_ASD3a, Slope_TD3a, Slope_ASD3b, Slope_TD3b, SD_intercept3, SD_visit3, Sigma3, ncol = 4)
```

```{r Posterior predictive update checks model 4}
# Prior-posterior update checks
posterior_m4 <- as_draws_df(df_post_4)
variables(posterior_m4)

# Intercept ASD update check
Intercept_ASD4 <- ggplot(posterior_m4) +
  geom_density(aes(prior_b_DiagnosisASD), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(b_DiagnosisASD), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab('Intercept ASD') +
  theme_classic() +
  labs(title = "Intercept ASD")

# Intercept TD update check
Intercept_TD4 <- ggplot(posterior_m4) +
  geom_density(aes(prior_b_DiagnosisTD), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(b_DiagnosisTD), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab('Intercept TD') +
  theme_classic() +
  labs(title = "Intercept TD")

# Slope ASD visit update check
Slope_ASD4a <- ggplot(posterior_m4) +
  geom_density(aes(`prior_b_DiagnosisASD:Visit`), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(`b_DiagnosisASD:Visit`), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab('Slope Visit ASD') +
  theme_classic() +
  labs(title = "Slope Visit ASD")

# Slope TD visit update check
Slope_TD4a <- ggplot(posterior_m4) +
  geom_density(aes(`prior_b_DiagnosisTD:Visit`), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(`b_DiagnosisTD:Visit`), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab('Slope Visit TD') +
  theme_classic() +
  labs(title = "Slope Visit TD")

# Slope ASD Socialization update check
Slope_ASD4b <- ggplot(posterior_m4) +
  geom_density(aes(`prior_b_DiagnosisASD:Socialization_1`), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(`b_DiagnosisASD:Socialization_1`), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab('Slope Socialization ASD') +
  theme_classic() +
  labs(title = "Slope Socialization ASD")

# Slope TD Socialization update check
Slope_TD4b <- ggplot(posterior_m4) +
  geom_density(aes(`prior_b_DiagnosisTD:Socialization_1`), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(`b_DiagnosisTD:Socialization_1`), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab('Slope Socialization TD') +
  theme_classic() +
  labs(title = "Slope Socialization TD")

# SD intercept update check
SD_intercept4 <- ggplot(posterior_m4) +
  geom_density(aes(prior_sd_ID__Intercept), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(sd_ID__Intercept), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab("SD intercept") +
  theme_classic() +
  labs(title = "SD intercept")

# SD visit update check
SD_visit4 <- ggplot(posterior_m4) +
  geom_density(aes(prior_sd_ID__Visit), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(sd_ID__Visit), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab("SD visit") +
  theme_classic() +
  labs(title = "SD visit")

# sigmaupdate check
Sigma4 <- ggplot(posterior_m4) +
  geom_density(aes(sigma), fill = "steelblue",
               color = "black", alpha = 0.6) +
  geom_density(aes(sigma), fill = "#FC4E07",
               color = "black", alpha = 0.6) +
  xlab("Sigma") +
  theme_classic() +
  labs(title = "Sigma")

grid.arrange(Intercept_ASD4, Intercept_TD4, Slope_ASD4a, Slope_TD4a, Slope_ASD4b, Slope_TD4b, SD_intercept4, SD_visit4, Sigma4, ncol = 4)
````

```{r Model comparison with Looic}
MLU_m1_posterior <- add_criterion(MLU_m1_posterior, criterion = "loo") #not included in model comparison
df_post_2 <- add_criterion(df_post_2, criterion = "loo")
df_post_3 <- add_criterion(df_post_3, criterion = "loo")
df_post_4 <- add_criterion(df_post_4, criterion = "loo")

loo_compare(df_post_2, df_post_3, df_post_4)

```




